---
title: "gsea_spearmans"
author: "Mehak Chopra"
date: "2025-06-18"
output: html_document
---
In the following, the analysis has been done using C2.GO only. 
GSEA analysis --- uploading all the genes with rho corr (estimates) , standard error and p values 
```{r}
library(dplyr)
library(readr)
library(purrr)

file_paths <- c("/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/all_genes/vsmc_se.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/all_genes/vsmcii_se.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/all_genes/fibroblastI_se.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/all_genes/fibroblastII_se.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/all_genes/Pericyte_se.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/all_genes/endothelialI_se.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/all_genes/endothelialII_se.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/all_genes/lym_endothelial_se.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/all_genes/lymphocyte_se.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/all_genes/macrophage_se.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/all_genes/mesothelial_se.csv",
               "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/all_genes/neuronal_se.csv")

celltype_names <- c("VSMCI", "VSMCII", "FibroblastI", "FibroblastII", "Pericyte", "EndothelialI", "EndothelialII", "Lymphatic_Endothelial", "Lymphocyte", "Macrophage", "Mesothelial", "Neuronal")

if (length(file_paths) != length(celltype_names)) {
  stop("The number of file paths does not match the number of celltype names.")
}

# 2. INITIALIZE YOUR LISTS
list_of_Estimates <- list()
list_of_ses <- list()
list_of_pvals <- list()

# 3. RUN THE LOOP
for (i in 1:length(file_paths)) {
  path <- file_paths[i]
  
  # --- KEY CHANGE: Get the custom name for this iteration ---
  current_celltype_name <- celltype_names[i]
  
  # Read the current file
  current_data <- readr::read_csv(path, show_col_types = FALSE) 
  
  # Process and store the Estimate data using the custom name
  list_of_Estimates[[i]] <- current_data %>%
    dplyr::select(Gene, Estimate) %>%
    # Use your custom name to create the column title
    dplyr::rename(!!paste0("Estimate_", current_celltype_name) := Estimate)

  # Process and store the standard error data
  list_of_ses[[i]] <- current_data %>%
    dplyr::select(Gene, SE) %>%
    # Use your custom name here as well
    dplyr::rename(!!paste0("se_", current_celltype_name) := SE)
  
  #Process and store the p values
  list_of_pvals[[i]] <- current_data %>%
    dplyr::select(Gene, pval) %>%
    dplyr::rename(!!paste0("pval_", current_celltype_name) := pval)
}

Estimate_df <- purrr::reduce(list_of_Estimates, full_join, by = "Gene")
se_df <- purrr::reduce(list_of_ses, full_join, by = "Gene")
pval_df <- purrr::reduce(list_of_pvals, full_join, by = "Gene") 

Estimate_final_df <- as.data.frame(Estimate_df)

# Now that it's a real data.frame, set the rownames
rownames(Estimate_final_df) <- Estimate_final_df$Gene

# Remove the now-redundant 'Gene' column
Estimate_final_df$Gene <- NULL

# Do the exact same process for the 'lfcSE' data
se_final_df <- as.data.frame(se_df)
rownames(se_final_df) <- se_final_df$Gene
se_final_df$Gene <- NULL

#pvalues
pval_final_df <- as.data.frame(pval_df)  # <-- New
rownames(pval_final_df) <- pval_final_df$Gene
pval_final_df$Gene <- NULL

# Step 4: Create the final list object 'resd'
resd_sp <- list(Estimate = Estimate_final_df, se = se_final_df, pval = pval_final_df)
```

In this, we will run the analysis using Hallmark genes, C2.CP (Canonical pathways), C5.GO (gene ontology), C5.HPO.We'll just change the genesets below if we are interested in any one. Now, as we are taking all the genes, there might be a small problem. This will give high ranking to positive correlated genes but will not consider negative correlations as it'll not consider them in analysis. That's the reason, performing analysis using positives and negatives separate is advisable.

#c2--canonical pathways
msigdb_df <- read.gmt("c2.cp.v2025.1.Hs.symbols.gmt")
#msigdb_sets <- split(msigdb_df$gene, msigdb_df$term)

#hallmark
#msigdb_h_df <- read.gmt("h.all.v2025.1.Hs.symbols.gmt")
#msigdb_h_sets <- split(msigdb_h_df$gene, msigdb_h_df$term)

#C8
#msigdb_c8_df <- read.gmt("c8.all.v2025.1.Hs.symbols.gmt")
#msigdb_c8_sets <- split(msigdb_c8_df$gene, msigdb_c8_df$term)

all the genes - GSEA 
```{r}
library(fgsea)
library(clusterProfiler)

msigdb_c5_df <- read.gmt("c5.go.v2025.1.Hs.symbols.gmt")
msigdb_c5_sets <- split(msigdb_c5_df$gene, msigdb_c5_df$term)

gene_stats_list <- list(
  VSMCI = resd_sp$Estimate$Estimate_VSMCI,
  VSMCII = resd_sp$Estimate$Estimate_VSMCII,
  FibroblastI = resd_sp$Estimate$Estimate_FibroblastI,
  FibroblastII = resd_sp$Estimate$Estimate_FibroblastII,
  Lymphatic_Endothelial = resd_sp$Estimate$Estimate_Lymphatic_Endothelial,
  Lymphocyte = resd_sp$Estimate$Estimate_Lymphocyte,
  Macrophage = resd_sp$Estimate$Estimate_Macrophage,
  Mesothelial = resd_sp$Estimate$Estimate_Mesothelial,
  Neuronal = resd_sp$Estimate$Estimate_Neuronal,
  Pericyte = resd_sp$Estimate$Estimate_Pericyte,
  EndothelialI = resd_sp$Estimate$Estimate_EndothelialI,
  EndothelialII = resd_sp$Estimate$Estimate_EndothelialII
                )

# Pre-process each vector in the list.
preproc_stats_list <- lapply(gene_stats_list, function(gene_stats_vector) {
  names(gene_stats_vector) <- rownames(resd_sp$Estimate)
  gene_stats_vector <- na.omit(gene_stats_vector)
  gene_stats_vector <- sort(gene_stats_vector, decreasing = TRUE)
  return(gene_stats_vector)
})


# --- Step 2: Run FGSEA with a Seed for Reproducibility ---
set.seed(42)
fgsea_results_list <- lapply(preproc_stats_list, function(ranked_genes) {
  fgsea(
    pathways = msigdb_c5_sets,
    stats = ranked_genes,
    minSize = 3,
    maxSize = 800
  )
})

# --- Step 3: Filter for significant pathways ---
fgsea_sig_list <- lapply(fgsea_results_list, function(res) {
  res[res$padj < 0.05 & !is.na(res$padj), ]
})
```


GSA - mremaR  -- all
```{r}
library(msigdbr)
library(clusterProfiler)

msigdb_c5_df <- read.gmt("c5.go.v2025.1.Hs.symbols.gmt")
msigdb_c5_sets <- split(msigdb_c5_df$gene, msigdb_c5_df$term)

full.gsa <- lapply(1:12, function(ct) {
  data <- data.frame(
    genes = rownames(resd_sp$Estimate),
    lfc = resd_sp$Estimate[, ct],
    lfcSE = resd_sp$se[, ct]
  )

  mremaR::REtest(data, log2(2), msigdb_c5_sets)
})

names(full.gsa) <- colnames(resd_sp$Estimate)[c(1:12)]
saveRDS(full.gsa, file = "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/mremaR_results/c5_go_all_genes_non_p.RDS")

#mremaR::REshine()
#full.gsa
```

If the results are already saved in RDS file
```{r}
mremaR_go <- readRDS("/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/mremaR_results/c5_go_all_genes_non_p.RDS")

VSMCI <- subset(mremaR_go$Estimate_VSMCI$results, P.value<0.05)
VSMCII <- subset(mremaR_go$Estimate_VSMCII$results, P.value<0.05)
fibroblastI <- subset(mremaR_go$Estimate_FibroblastI$results, P.value<0.05)
fibroblastII <- subset(mremaR_go$Estimate_FibroblastII$results, P.value<0.05)
Pericyte <- subset(mremaR_go$Estimate_Pericyte$results, P.value<0.05)
EndothelialI <- subset(mremaR_go$Estimate_EndothelialI$results, P.value<0.05)
EndothelialII <- subset(mremaR_go$Estimate_EndothelialII$results, P.value<0.05)
Lymphocyte <- subset(mremaR_go$Estimate_Lymphocyte$results, P.value<0.05)
Macrophage <- subset(mremaR_go$Estimate_Macrophage$results, P.value<0.05)
Mesothelial <- subset(mremaR_go$Estimate_VSMCI$results, P.value<0.05)
Lymphatic_Endothelial <- subset(mremaR_go$Estimate_VSMCI$results, P.value<0.05)
Neuronal <- subset(mremaR_go$Estimate_VSMCI$results, P.value<0.05)

all_mremaR_go <- list(VSMCI, VSMCII, fibroblastI, fibroblastII, Pericyte, EndothelialI, EndothelialII, Lymphocyte, Macrophage, Mesothelial, Lymphatic_Endothelial, Neuronal)

names(all_mremaR_go) <- c("VSMCI", "VSMCII", "Fibroblast-I", "Fibroblast-II", "Pericyte", "Endothelial-I", "Endothelial-II", "Lymphocyte", "Macrophage", "Mesothelial", "Lymphatic Endothelial", "Neuronal")
```

just the significant genes -- fix the names so that it doesnt overlap anywhere
```{r}
file_paths <- c("/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/vsmc_se_significant.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/vsmcii_se_significant.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/fibroblastI_se_significant.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/fibroblastII_se_significant.csv",
              "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/Pericyte_se_significant.csv",
              "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/endothelialI_se_significant.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/endothelialII_se_significant.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/lym_endothelial_se_significant.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/lymphocyte_se_significant.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/macrophage_se_significant.csv",
                "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/mesothelial_se_significant.csv",
            "/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/hires_age_associations/gsea_spearman/neuronal_se_significant.csv")

celltype_names <- c("VSMCI", "VSMCII", "FibroblastI", "FibroblastII", "Pericyte", "EndothelialI", "EndothelialII", "Lymphatic_Endothelial", "Lymphocyte", "Macrophage", "Mesothelial", "Neuronal")

if (length(file_paths) != length(celltype_names)) {
  stop("The number of file paths does not match the number of celltype names.")
}

# 2. INITIALIZE YOUR LISTS
list_of_Estimates <- list()
list_of_ses <- list()
list_of_pvals <- list()

# 3. RUN THE LOOP
for (i in 1:length(file_paths)) {
  path <- file_paths[i]
  
  # --- KEY CHANGE: Get the custom name for this iteration ---
  current_celltype_name <- celltype_names[i]
  
  # Read the current file
  current_data <- readr::read_csv(path, show_col_types = FALSE) 
  
  # Process and store the Estimate data using the custom name
  list_of_Estimates[[i]] <- current_data %>%
    dplyr::select(Gene, Estimate) %>%
    # Use your custom name to create the column title
    dplyr::rename(!!paste0("Estimate_", current_celltype_name) := Estimate)

  # Process and store the standard error data
  list_of_ses[[i]] <- current_data %>%
    dplyr::select(Gene, SE) %>%
    # Use your custom name here as well
    dplyr::rename(!!paste0("se_", current_celltype_name) := SE)
  
  #Process and store the p values
  list_of_pvals[[i]] <- current_data %>%
    dplyr::select(Gene, pval) %>%
    dplyr::rename(!!paste0("pval_", current_celltype_name) := pval)
}

Estimate_df <- purrr::reduce(list_of_Estimates, full_join, by = "Gene")
se_df <- purrr::reduce(list_of_ses, full_join, by = "Gene")
pval_df <- purrr::reduce(list_of_pvals, full_join, by = "Gene") 

Estimate_final_df <- as.data.frame(Estimate_df)

# Now that it's a real data.frame, set the rownames
rownames(Estimate_final_df) <- Estimate_final_df$Gene

# Remove the now-redundant 'Gene' column
Estimate_final_df$Gene <- NULL

# Do the exact same process for the 'lfcSE' data
se_final_df <- as.data.frame(se_df)
rownames(se_final_df) <- se_final_df$Gene
se_final_df$Gene <- NULL

#pvalues
pval_final_df <- as.data.frame(pval_df)  # <-- New
rownames(pval_final_df) <- pval_final_df$Gene
pval_final_df$Gene <- NULL

# Step 4: Create the final list object 'resd'
resd_sp_sign <- list(Estimate = Estimate_final_df, se = se_final_df, pval = pval_final_df)
```

ORA - fgsea 
```{r}
library(fgsea)

# Load pathways
pathways_gmt_file <- "c5.go.v2025.1.Hs.symbols.gmt"
msigdb_c5_sets <- gmtPathways(pathways_gmt_file)

# Define the BACKGROUND gene list from the full results table.
# This will be the same for all 12 analyses.
experiment_columns <- colnames(resd_sp$Estimate)

all_ora_results <- lapply(experiment_columns, function(col_name) {
  
  message(paste("--- Running ORA for experiment:", col_name, "---"))
  
  # A. Define the FOREGROUND genes using your exact logic.
  foreground_values <- resd_sp_sign$Estimate[[col_name]]
  names(foreground_values) <- rownames(resd_sp_sign$Estimate)
  foreground_values <- na.omit(foreground_values)
  foreground_genes <- names(foreground_values)
  
  # B. Define the BACKGROUND genes using your exact logic.
  background_values <- resd_sp$Estimate[[col_name]]
  names(background_values) <- rownames(resd_sp$Estimate)
  background_values <- na.omit(background_values)
  background_genes <- names(background_values)
  
  # Safety check and progress message
  if (length(foreground_genes) == 0) {
    message("   No significant (non-NA) genes found for this experiment. Skipping.")
    return(NULL) # Return NULL so the loop can continue
  }
  
  message(paste("   Found", length(foreground_genes), "foreground genes"))
  message(paste("   Using a background of", length(background_genes), "genes for this specific experiment."))
  
  # C. Run the Over-Representation Analysis
  foraRes <- fora(
    genes = foreground_genes,
    universe = background_genes,
    pathways = msigdb_c5_sets,
    minSize = 3,
    maxSize = 800
  )
    foraRes <- foraRes[foraRes$padj < 0.05, ]
  
  # E. Add the overlapping genes column for the remaining enriched pathways.
  if (nrow(foraRes) > 0) {
    overlapping_genes_list <- sapply(foraRes$pathway, function(pathway_name) {
      genes_in_pathway <- msigdb_c5_sets[[pathway_name]]
      overlapping_genes <- intersect(foreground_genes, genes_in_pathway)
      paste(overlapping_genes, collapse = ", ")
    })
    foraRes$foreground_genes_in_pathway <- overlapping_genes_list
  }
  
  # F. Return the final, filtered table for this experiment.
  return(foraRes)
})

names(all_ora_results) <- experiment_columns
```

```{r}
common_pathways_VSMCI <- unlist(Reduce(intersect, list(fgsea_sig_list$VSMCI$pathway, all_mremaR_go$VSMCI$Gene.Set, all_ora_results$Estimate_VSMCI$pathway)))

common_pathways_VSMCII <- unlist(Reduce(intersect, list(fgsea_sig_list$VSMCII$pathway, all_mremaR_go$VSMCII$Gene.Set, all_ora_results$Estimate_VSMCII$pathway)))

common_pathways_fibroblastI <- unlist(Reduce(intersect, list(fgsea_sig_list$FibroblastI$pathway, all_mremaR_go$`Fibroblast-I`$Gene.Set, all_ora_results$Estimate_FibroblastI$pathway)))

common_pathways_fibroblastII <- unlist(Reduce(intersect, list(fgsea_sig_list$FibroblastII$pathway, all_mremaR_go$`Fibroblast-II`$Gene.Set, all_ora_results$Estimate_FibroblastII$pathway)))

common_pathways_EndothelialI <- unlist(Reduce(intersect, list(fgsea_sig_list$EndothelialI$pathway, all_mremaR_go$`Endothelial-I`$Gene.Set, all_ora_results$Estimate_EndothelialI$pathway)))

common_pathways_EndothelialII <- unlist(Reduce(intersect, list(fgsea_sig_list$EndothelialII$pathway, all_mremaR_go$`Endothelial-II`$Gene.Set, all_ora_results$Estimate_EndothelialII$pathway)))

common_pathways_lymphocyte <- unlist(Reduce(intersect, list(fgsea_sig_list$Lymphocyte$pathway, all_mremaR_go$Lymphocyte$Gene.Set, all_ora_results$Estimate_Lymphocyte$pathway)))

common_pathways_macrophage <- unlist(Reduce(intersect, list(fgsea_sig_list$Macrophage$pathway, all_mremaR_go$Macrophage$Gene.Set, all_ora_results$Estimate_Macrophage$pathway)))

common_pathways_mesothelial <- unlist(Reduce(intersect, list(fgsea_sig_list$Mesothelial$pathway, all_mremaR_go$Mesothelial$Gene.Set, all_ora_results$Estimate_Mesothelial$pathway)))

common_pathways_pericyte <- unlist(Reduce(intersect, list(fgsea_sig_list$Pericyte$pathway, all_mremaR_go$Pericyte$Gene.Set, all_ora_results$Estimate_Pericyte$pathway)))

common_pathways_neuronal <- unlist(Reduce(intersect, list(fgsea_sig_list$Neuronal$pathway, all_mremaR_go$Neuronal$Gene.Set, all_ora_results$Estimate_Neuronal$pathway)))

common_pathways_VSMCI
common_pathways_VSMCII
common_pathways_fibroblastI
common_pathways_fibroblastII
common_pathways_EndothelialI
common_pathways_EndothelialII
common_pathways_lymphocyte
common_pathways_macrophage
common_pathways_mesothelial
common_pathways_pericyte
common_pathways_neuronal

#extracting fgsea values 
common_VSMCI <- fgsea_sig_list$VSMCI[fgsea_sig_list$VSMCI$pathway %in% common_pathways_VSMCI, ]
common_VSMCI$celltype <- "VSMCI"
```

Because we get only 5 pathways common in VSMCI
```{r}
ggplot(common_VSMCI, aes(x = NES, y = reorder(pathway, NES), color = padj)) +
  geom_point(size = 3) +
  #geom_density() +
  scale_color_gradient(low = "blue", high = "red", trans = "reverse", name = "Adj. P-value") +
  theme_minimal() +
  labs(
    title = "GSEA Pathways by NES and Significance in VSMCI",
    x = "Normalized Enrichment Score (NES)",
    y = "Pathway"
  )
```

common between fGSEA and mremaR 
```{r}
# Find common pathways
common_pathways_VSMCI <- intersect(fgsea_sig_list$VSMCI$pathway, all_mremaR_go$VSMCI$Gene.Set)
common_VSMCI <- fgsea_sig_list$VSMCI[fgsea_sig_list$VSMCI$pathway %in% common_pathways_VSMCI, ]
common_VSMCI$celltype <- "VSMCI"

common_pathways_VSMCII <- intersect(fgsea_sig_list$VSMCII$pathway, all_mremaR_go$VSMCII$Gene.Set)
common_VSMCII <- fgsea_sig_list$VSMCII[fgsea_sig_list$VSMCII$pathway %in% common_pathways_VSMCII, ]
common_VSMCII$celltype <- "VSMCII"

common_pathways_fibroblastI <- intersect(fgsea_sig_list$FibroblastI$pathway, all_mremaR_go$`Fibroblast-I`$Gene.Set)
common_fibroblastI <- fgsea_sig_list$FibroblastI[fgsea_sig_list$FibroblastI$pathway %in% common_pathways_fibroblastI, ]
common_fibroblastI$celltype <- "Fibroblast-I"

common_pathways_fibroblastII <- intersect(fgsea_sig_list$FibroblastII$pathway, all_mremaR_go$`Fibroblast-II`$Gene.Set)
common_fibroblastII <- fgsea_sig_list$FibroblastII[fgsea_sig_list$FibroblastII$pathway %in% common_pathways_fibroblastII, ]
common_fibroblastII$celltype <- "Fibroblast-II"

common_pathways_EndothelialI <- intersect(fgsea_sig_list$EndothelialI$pathway, all_mremaR_go$`Endothelial-I`$Gene.Set)
common_EndothelialI <- fgsea_sig_list$EndothelialI[fgsea_sig_list$EndothelialI$pathway %in% common_pathways_EndothelialI, ]
common_EndothelialI$celltype <- "Endothelial-I"

common_pathways_EndothelialII <- intersect(fgsea_sig_list$EndothelialII$pathway, all_mremaR_go$`Endothelial-II`$Gene.Set)
common_EndothelialII <- fgsea_sig_list$EndothelialII[fgsea_sig_list$EndothelialII$pathway %in% common_pathways_EndothelialII, ]
common_EndothelialII$celltype <- "Endothelial-II"

common_pathways_lymphocyte <- intersect(fgsea_sig_list$Lymphocyte$pathway, all_mremaR_go$Lymphocyte$Gene.Set)
common_lymphocyte <- fgsea_sig_list$Lymphocyte[fgsea_sig_list$Lymphocyte$pathway %in% common_pathways_lymphocyte, ]
common_lymphocyte$celltype <- "Lymphocyte"

common_pathways_Macrophage <- intersect(fgsea_sig_list$Macrophage$pathway, all_mremaR_go$Macrophage$Gene.Set)
common_Macrophage <- fgsea_sig_list$Macrophage[fgsea_sig_list$Macrophage$pathway %in% common_pathways_Macrophage, ]
common_Macrophage$celltype <- "Macrophage"

common_pathways_Mesothelial <- intersect(fgsea_sig_list$Mesothelial$pathway, all_mremaR_go$Mesothelial$Gene.Set)
common_Mesothelial <- fgsea_sig_list$Mesothelial[fgsea_sig_list$Mesothelial$pathway %in% common_pathways_Mesothelial, ]
common_Mesothelial$celltype <- "Mesothelial"

common_pathways_Pericyte <- intersect(fgsea_sig_list$Pericyte$pathway, all_mremaR_go$Pericyte$Gene.Set)
common_Pericyte <- fgsea_sig_list$Pericyte[fgsea_sig_list$Pericyte$pathway %in% common_pathways_Pericyte, ]
common_Pericyte$celltype <- "Pericyte"

common_pathways_Neuronal <- intersect(fgsea_sig_list$Neuronal$pathway, all_mremaR_go$Neuronal$Gene.Set)
common_Neuronal<- fgsea_sig_list$Neuronal[fgsea_sig_list$Neuronal$pathway %in% common_pathways_Neuronal, ]
common_Neuronal$celltype <- "Neuronal"

all_common <- rbind(common_VSMCI, common_VSMCII, common_fibroblastI, common_fibroblastII, common_EndothelialI, common_EndothelialII, common_lymphocyte, common_Macrophage, common_Mesothelial, common_Pericyte, common_Neuronal)

library(ggplot2)
gg <- ggplot(all_common, 
       aes(x = celltype, y = pathway, colour = padj, size = size)) +
      geom_point() +
      theme_minimal() +
     theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.5))
gg
#ggsave("/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/genesets_common_fgsea_mremaR.png", gg, height = 10, width = 15, dpi = 300)
```

```{r}
library(ggtext)

celltype_cols <- c(VSMCI="#333333", VSMCII="#B2EF9B", "Fibroblast-I"="#3A7219", Lymphocyte="#1B360C")

gradient_cols <- colorRampPalette(c('#190c3E', '#87216B', '#E44C30', '#F7D340'))

y_labs <- paste0("<span style='color:", celltype_cols, "'>", names(celltype_cols), "</span>")
#y_labs <- paste0(names(celltype_cols))
#head(y_labs)

xxx<- ggplot(all_common, 
       aes(y = celltype, x = pathway, colour = padj, size = size)) +
      geom_point() +
    #  scale_x_discrete(limits = gene_order) +
  scale_y_discrete(limits = rev(names(celltype_cols)), labels = rev(y_labs)) + coord_flip() +
  scale_colour_gradientn(colours = gradient_cols(100)) +
  scale_size(range = c(0.5,4)) +
 # facet_wrap(~celltype) +
#  scale_size(labels = scales::percent, range = c(0.5, 4)) +
  labs(y = "celltypes", x = "pathways", colour = 'P adj', size = 'number of genes') +
  guides(colour = guide_colourbar(order = 1, barlength = unit(6, 'lines')), size = guide_legend(order = 2)) +
  theme_linedraw(base_size = 14) +
  theme(axis.text.x = element_markdown(angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_markdown(face = 'bold'),
        legend.title = element_text(size = 12, vjust = 0.5),
        legend.box.margin = margin(t = 45))
xxx

#ggsave("/home/mchopra/Documents/PhD-Year1/deconvolution/Deconvolution_results/results/with_batchs/hires/genesets_common_fgsea_mremaR_colour.png", xxx, height = 15, width = 20, dpi = 300)
```
